---
title: "Statistical model that predicts shoe size given person tallness"
author: "Rafal Szota"
date: "November 22, 2015"
output: ioslides_presentation
---

## Model overview

This application is based on very simple linear model that correlates person tallness with their shoe size.

- The input person tallness is done in centimeters
- Resulting shoe size is predicted according to European standards  
- Prediction uses linear regression fitted to 100 examples
- Future model extensions could consider sex as predictor to increase accuracy

## Sample Data 1/2

Data for the model is availible in GitHub. <http://http://rafalszota.github.io/>.

```{r, echo=FALSE}
data <- read.csv("data.txt") 
```

```{r}
summary(data)
```


## Sample Data 2/2

```{r, echo=FALSE}
plot(data$height, data$footsize,
     ylab="Shoe size (EU)",
     xlab="Tallness in centimeters",
     main="Correlation between tallness and shoe size")

```


## Prediction model


The prediction model uses regression line with the following paremeters fitted:

```{r}
lm(data$footsize ~ data$height)
```
Predictions are rounded to the nearest half using the following formula:
```{r, results=F, eval=FALSE}
0.5 * round(predictedValue * 0.5)
```